---
title: "Natural Lacerate Totals"
author: "Erick White"
date: "2023-07-10"
output: pdf_document
---
#looking at the total lacerate data over time throughout the whole collection
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyr)
library(dplyr)
library(scales)
library(emmeans)

#Total natural lacerate data

View(Natural_Lacerate_collection_log)
# Convert the "Date" column to a proper date format
Natural_Lacerate_collection_log$Date <- as.Date(Natural_Lacerate_collection_log$Date, format = "%d-%b")

# Define the dates you want to show on the x-axis
selected_dates <- c("15-May", "29-May", "12-Jun", "26-Jun", "10-Jul")
selected_dates <- as.Date(selected_dates, format = "%d-%b")

# Create a scatter plot with selected dates on the x-axis
ggplot(Natural_Lacerate_collection_log, aes(x = Date, y = Total, color = `Symbiotic State`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, aes(group = `Symbiotic State`)) +
  labs(x = "Date", y = "Total Lacerates Collected") +
  scale_x_date(breaks = selected_dates, labels = date_format("%d-%b")) +
  scale_color_manual(values = c("Sym" = "orange", "Apo" = "darkgrey")) +
  theme_minimal()


#Statistics for total lacerate data
Lacerates<-lm(Total~Date, data=Natural_Lacerate_collection_log)
summary(Lacerates)

# Create separate linear regression models for "Sym" and "Apo"
lm_sym <- lm(Total ~ Date, data = subset(Natural_Lacerate_collection_log, `Symbiotic State` == "Sym"))
lm_apo <- lm(Total ~ Date, data = subset(Natural_Lacerate_collection_log, `Symbiotic State` == "Apo"))

# Print the regression summary for "Sym"
summary(lm_sym)

# Print the regression summary for "Apo"
summary(lm_apo)

# Perform hypothesis tests on the coefficients
coef_test <- coef(summary(lm_sym))
coef_test2 <- coef(summary(lm_apo))

p_value <- coef_test["Date", "Pr(>|t|)"]
p_value2 <- coef_test2["Date", "Pr(>|t|)"]
# Display the p-value
p_value
p_value2
```
`
#looking at how the different bins of ages affected the number of lacerates collected i.e. is it better to wait a few days or collect every day

```{r}
filtered_data <- Natural_Lacerate_collection_log %>%
  filter(`Age` %in% c("0-1", "0-2", "0-3")) %>%
  filter(!`Lacerates` %in% boxplot.stats(`Lacerates`)$out)

ggplot(filtered_data, aes(x = `Age`, y = `Lacerates`, fill = `Symbiotic State`)) +
  geom_boxplot(outlier.shape = NA) +
  labs(x = "Lacerate Age", y = "Lacerate Number", fill = "Symbiotic State") +
  scale_x_discrete(labels = function(x) gsub("-", "-", x)) +
  scale_fill_manual(values = c("Sym" = "orange", "Apo" = "darkgrey")) +
  theme_minimal()

# Perform ANOVA
anova_result <- aov(`Lacerates` ~ `Age` * `Symbiotic State`, data = filtered_data)

# Compute pairwise comparisons with emmeans
emmeans_result <- emmeans(anova_result, ~ `Age` * `Symbiotic State`)

# Perform Tukey adjustment for multiple comparisons
tukey_result <- contrast(emmeans_result, method = "tukey")

# Print ANOVA results
print(summary(anova_result))

# Print Tukey test results
print(tukey_result)

```

there was only one time point for the 0-4 and 0-5 group, so those results are outliers and may not reflect actual patterns. I decided to remove those groups.

#bar graph figure
```{r}
#Barplot grapj
averages <- Natural_Lacerate_collection_log %>%
  group_by(`Lacerate Age`, `Symbiotic State`) %>%
  summarize(Average = mean(`Lacerate Number`))
averages

ggplot(averages, aes(x = `Lacerate Age`, y = `Average`, fill = `Symbiotic State`)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Age of Lacerates in Days", y = "Average Number of Lacerates Collected") +
  scale_fill_manual(values = c("Sym" = "orange", "Apo" = "darkgrey")) +
  theme_minimal()

```

